### 2020-04-22 intro.tex
intro tweak

\begin{figure*}[tb]
  \centering
  \includegraphics[width=\textwidth]{figs/chronology001.pdf}
  \caption{Volume of GPU swaps detected on Titan at inventories. Our
    analysis begins with units in place by 2014. A break-in period
    before 2014 involved working out mechanical assembly issues at
    PCIe connectors.}
  \label{fig:chronology}
\end{figure*}
Figure~\ref{fig:chronology} illustrates the chronology of the rework
cycles and stable periods as indicated by the number of GPU changes at
periodic inventories.
The first two rework cycles before 2014 can be characterized as a
break-in period on a new system that is first of its kind and pushes
many technological boundaries, in this case mainly regarding blade
mechanical assembly. The resulting field engineering generated high
GPU swap counts. Late 2013 begins a very long period of stable
operation with very few issues until about mid 2016, when GPU failures
begin to rise. This results in the final rework cycle, replacing
11,000 GPUs throughout much of 2017.

The failures were traced to a failing resistor on the circuitboard
(not the GPU chip iself). Throughout this paper, when we refer to
replacing the GPU, we refer to the entire circuitboard along with its
GPU chip. While the 2016 resistor failures were a surprise, there is a
known phenomenon~\cite{referenceneeded} that describes a two phase
process that leads to failure. It also explains why there was a very
stable period preceding the failures. There is specific decay
associated with the electronic components where, as these products
age, they will eventually fail thresholds for guard band margin and
other things. In a lot of cases, they will just fail as
well. \fix{Give a reference and expand?}

Our analysis focuses on GPUs that were installed in the second rework
cycle and later, essentialy units installed near the beginning of 2014
and later. We will make the data as well as our analysis codes
producing the graphics available in a public repository (see
\fix{reference?}).  The data used in this study, while only a 7 MB
file, represents over 100,000 years of operation and may be the
largest publicly available GPU reliability data set.

>>>>

\begin{figure*}[tb]
  \centering
  \includegraphics[width=\textwidth]{figs/chronology001.pdf}
  \caption{Volume of GPU swaps detected on Titan at inventories. Our
    analysis begins with units in place by 2014. A break-in period
    before 2014 involved working out mechanical assembly issues at
    PCIe connectors.}
  \label{fig:chronology}
\end{figure*}
Figure~\ref{fig:chronology} illustrates the chronology of the rework
cycles and stable periods as indicated by the number of GPU changes at
periodic inventories.  The first two rework cycles before 2014 can be
characterized as a break-in period on a new system that is first of
its kind and pushes many technological boundaries, in this case mainly
regarding blade mechanical assembly. The resulting field engineering
and data collection changes generated high GPU swap counts before
2014. Late 2013 begins a very long period of high reliability and
stable operation with very few issues until about mid 2016, when GPU
failures begin to rise. This results in the final rework cycle,
replacing 11,000 GPUs throughout much of 2017. The peak at the end in
the figure is an artifact of turning off or ``removing'' all the
units.

Finding the root cause for the 2016 failures took a great deal of
testing and involvement of materials science and microscopy
researchers. The failures were traced to a resistor on the GPU
circuit board (not the GPU chip iself). Throughout this paper, when we
refer to replacing the GPU, we refer to the entire circuitboard along
with its GPU chip.

Our analysis focuses on GPU boards that were installed in the second
rework cycle and later, essentially units installed near the beginning
of 2014 and later. We will make the data as well as our analysis codes
producing the graphics available in a public repository (see,
\cite[repository]{repositoryreference}).  While the data is only a 7 MB
file, due to the size of Titan it represents over 100,000 years of GPU
operation and may be the largest publicly available GPU reliability
data set. Our hope is that others will use it for reliability research
and go beyond the results presented here.
